---
title: "Lesson1GIT"
author: "Jadyn Abeyta"
date: "2024-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set up
```{r}
packageLoad <-
  function(x) {
    for (i in 1:length(x)) {
      if (!x[i] %in% installed.packages()) {
        install.packages(x[i])
      }
      library(x[i], character.only = TRUE)
    }
  }

# create a string of package names
packages <- c('tidyverse',
              'palmerpenguins',
              'lterdatasampler')

packageLoad(packages)

```
assignment
```{r}
data("knz_bison")
```
Create a new column in the bison dataset that calculates the individual's age at the time of observation.
```{r}
str(knz_bison)
head(knz_bison)
bison_age <- knz_bison %>%
  mutate(age_at_observation = rec_year - animal_yob)

```
2.  Write a function that will first filter the bison dataset based on any user-specified weight threshold (returning the filtered dataset).
```{r}
filter_bison_by_weight <- function(knz_bison, weight_threshhold) {
  if (!"animal_weight" %in% names(knz_bison)) {
    stop("The dataset does not contain a 'weight' column.")
  }
  
  filtered_bison <- knz_bison %>%
    filter(animal_weight >= weight_threshhold)
  
  return(filtered_bison)
}

filtered_bison <- filter_bison_by_weight(knz_bison, 500)

head(filtered_bison)

```
3.  Create a boxplot showing weight distribution between male and female bison, and then conduct a t-test to see if weight is significantly difference between males and females.
```{r}
library(ggplot2)
library(dplyr)

# Ensure that 'Weight' and 'Sex' columns exist in the dataset
# Adjust the column names as per your dataset structure
names(knz_bison)  # Check for exact column names (e.g., weight and sex)

# Create a boxplot for weight distribution between males and females
ggplot(knz_bison, aes(x = animal_sex, y = animal_weight, fill = animal_sex)) +
  geom_boxplot() +
  labs(title = "Weight Distribution by Sex in Bison",
       x = "Sex",
       y = "Weight (kg)") +
  theme_minimal()

# Conduct a t-test to compare weight between males and females
# Ensure that Sex is coded correctly (e.g., "Male" and "Female")
t_test_result <- t.test(animal_weight ~ animal_sex ,data = knz_bison)

# Print the t-test result
print(t_test_result)

```

